---
title: "Dataset assembly"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dataset assembly}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette documents how the `cqtkit_data_*` and `cqtkit_data_bl_*`
datasets shipped with the package were assembled from the raw trial
data, and shows how cqtkit's helper functions support C-QT dataset
assembly more generally.

The raw source file is **not** redistributed with cqtkit, so the code
chunks below are shown for reference and are not evaluated when the
vignette is built. To reproduce the datasets, point `read.csv()` at
your own copy of the source file.

## Source

The datasets in this package — the four per-drug `cqtkit_data_*` files
and the four `cqtkit_data_bl_*` baseline files — are derived from the
FDA ECGRDVQ trial reported in Johannesen et al., *CPT* 2014
(<https://doi.org/10.1038/clpt.2014.155>). The assembly uses two source
files:

- `SCR-002.Clinical.Data.csv` — ECG-level measurements
- `SCR-002.Clinical.Data.Description.txt` — column dictionary

Per the dictionary, `EXTRT` carries the human-readable treatment name
(`Ranolazine`, `Dofetilide`, `Quinidine Sulph`, `Verapamil HCL`,
`Placebo`), and `BASELINE` is a Y/N flag identifying pre-dose ECGs.

## Setup

```{r, message = FALSE}
library(dplyr)
library(cqtkit)

raw <- read.csv("SCR-002.Clinical.Data.csv", stringsAsFactors = FALSE)
```

## Baseline datasets (`cqtkit_data_bl_*`)

For each drug, keep pre-dose baseline rows (`BASELINE == "Y"`,
`TPT == -0.5`) for both the drug arm and the Placebo arm, and select
the ECG columns plus subject-level demographics (`SEX`, `AGE`, `HGHT`,
`WGHT`, `RACE`, `ETHNIC`, `VISIT`). Heart rate is derived from RR
(`60000 / RR`, ms to bpm). Rows with missing `QT` are dropped. `TRTG`
is a two-level factor (`Placebo | <drug>`), with both levels populated.
Including the Placebo arm makes each baseline file self-contained for
the corresponding two-arm "study", so it can be passed directly as
`bl_data` to `compute_hrblm()` / `compute_qtcbblm()` /
`compute_qtcfblm()` to reproduce that study's population baseline
means.

```{r}
build_bl <- function(drug_label, raw) {
  raw |>
    filter(
      BASELINE == "Y",
      EXTRT %in% c("Placebo", drug_label),
      !is.na(QT)
    ) |>
    mutate(TRTG = factor(EXTRT, levels = c("Placebo", drug_label))) |>
    compute_hr(rrbl_col = NULL) |>
    compute_qtcb_qtcf(qtbl_col = NULL, rrbl_col = NULL) |>
    select(
      ID = RANDID,
      SEX, AGE, HGHT, WGHT, RACE, ETHNIC, VISIT,
      TRTG,
      DOSE = EXDOSE,
      DOSEU = EXDOSU,
      TPT, RR, QT, HR, QTCB, QTCF
    ) |>
    tibble::as_tibble()
}

cqtkit_data_bl_verapamil <- build_bl("Verapamil HCL", raw)
cqtkit_data_bl_dofetilide <- build_bl("Dofetilide", raw)
cqtkit_data_bl_quinidine <- build_bl("Quinidine Sulph", raw)
cqtkit_data_bl_ranolazine <- build_bl("Ranolazine", raw)
```

A quick audit of the assembled baselines:

```{r}
bls <- list(
  verapamil = cqtkit_data_bl_verapamil,
  dofetilide = cqtkit_data_bl_dofetilide,
  quinidine = cqtkit_data_bl_quinidine,
  ranolazine = cqtkit_data_bl_ranolazine
)

for (slug in names(bls)) {
  d <- bls[[slug]]
  cat(sprintf(
    "%-10s  rows = %2d   subjects = %2d   TRTG = %s\n",
    slug, nrow(d), dplyr::n_distinct(d$ID),
    paste(levels(d$TRTG), collapse = " | ")
  ))
}
```

The baseline datasets are written with `usethis::use_data()`

## Analysis datasets (`cqtkit_data_*`)

Each analysis dataset combines one drug arm with the placebo arm of the
same crossover. Construction rules:

- **Triplicate averaging**: post-dose rows are averaged within each
  `(ID, EXTRT, TPT)` group. A group is dropped entirely if any of its
  three QT values is `NA` (no NA imputation).
- **Per-subject-per-arm baseline**: `RRBL`, `QTBL`, `HRBL`, `QTCBBL`,
  `QTCFBL` come from the per-subject-per-arm averaged baseline rows.
- **Derived per-row metrics computed pre-average**: `HR`, `QTCB`,
  `QTCF` are computed per raw ECG row (`60000/RR`, `QT/sqrt(RR/1000)`,
  `QT/(RR/1000)^(1/3)`) and then averaged across the triplicate.
  Averaging RR and QT first and then applying the corrections gives
  different values because the corrections are nonlinear in RR.
- **Population baseline means (`HRBLM`, `QTCBBLM`, `QTCFBLM`)** are
  per-drug constants computed across the baseline rows of the two arms
  in that drug's study (drug arm + Placebo arm), via `compute_hrblm()` /
  `compute_qtcbblm()` / `compute_qtcfblm()` with `by = c(ID, TRTG)` —
  per-subject-per-period averaging then population mean, per the white
  paper. Each drug's `*BLM` values differ from the other drugs'.
- **Deltas**: the baseline-from-mean deltas (`deltaHRBL`/`deltaQTCBBL`/
  `deltaQTCFBL`) and the row-level deltas
  (`deltaQTCB`/`deltaQTCF`/`deltaHR`/`deltaQT`/`deltaRR`).

The tail of `build_main()` below adds the population means
(`compute_hrblm()` / `compute_qtcbblm()` / `compute_qtcfblm()`), the
baseline-from-mean deltas (`compute_delta_hrblm()` /
`compute_delta_qtcbblm()` / `compute_delta_qtcfblm()`), and the
row-level deltas (`compute_deltas()`).

```{r}
build_main <- function(drug_label, raw, bl_data) {
  bl <- bl_data |>
    group_by(ID, TRTG) |>
    summarise(
      RRBL = mean(RR),
      QTBL = mean(QT),
      HRBL = mean(HR),
      QTCBBL = mean(QTCB),
      QTCFBL = mean(QTCF),
      .groups = "drop"
    )

  post <- raw |>
    filter(BASELINE == "N", EXTRT %in% c("Placebo", drug_label)) |>
    compute_qtcb_qtcf(qtbl_col = NULL, rrbl_col = NULL) |>
    compute_hr(rrbl_col = NULL) |>
    group_by(RANDID, EXTRT, EXDOSE, EXDOSU, TPT) |>
    filter(!any(is.na(QT))) |>
    summarise(
      RR = mean(RR),
      QT = mean(QT),
      HR = mean(HR),
      QTCB = mean(QTCB),
      QTCF = mean(QTCF),
      CONC = mean(PCSTRESN, na.rm = TRUE),
      CONCU = dplyr::first(PCSTRESU),
      .groups = "drop"
    )

  drug_dose <- post |> filter(EXTRT == drug_label) |> dplyr::slice(1) |> pull(EXDOSE)
  drug_unit <- post |> filter(EXTRT == drug_label) |> dplyr::slice(1) |> pull(EXDOSU)

  post |>
    inner_join(bl, by = c("RANDID" = "ID", "EXTRT" = "TRTG")) |>
    mutate(
      TRTG = factor(EXTRT, levels = c("Placebo", drug_label)),
      DOSE = ifelse(EXTRT == "Placebo", 0, drug_dose),
      DOSEU = ifelse(EXTRT == "Placebo", "mg", drug_unit),
      DOSEF = factor(
        paste(DOSE, DOSEU),
        levels = c("0 mg", paste(drug_dose, drug_unit))
      ),
      NTLD = TPT,
      TAFD = paste(TPT, "HR"),
      CONC = ifelse(EXTRT == "Placebo", 0, CONC),
      CONCU = ifelse(EXTRT == "Placebo", NA_character_, CONCU)
    ) |>
    compute_hrblm(bl_data, by = c(ID, TRTG)) |>
    compute_qtcbblm(bl_data, by = c(ID, TRTG)) |>
    compute_qtcfblm(bl_data, by = c(ID, TRTG)) |>
    compute_delta_hrblm() |>
    compute_delta_qtcbblm() |>
    compute_delta_qtcfblm() |>
    compute_deltas() |>
    select(
      ID = RANDID, TRTG, DOSE, DOSEU, DOSEF, NTLD, TAFD, CONC, CONCU,
      RR, RRBL, deltaRR,
      HR, HRBL, HRBLM, deltaHRBL, deltaHR,
      QT, QTBL, deltaQT,
      QTCB, QTCBBL, QTCBBLM, deltaQTCBBL, deltaQTCB,
      QTCF, QTCFBL, QTCFBLM, deltaQTCFBL, deltaQTCF
    ) |>
    tibble::as_tibble() |>
    arrange(TRTG, NTLD, ID)
}

cqtkit_data_verapamil <- build_main("Verapamil HCL", raw, cqtkit_data_bl_verapamil)
cqtkit_data_dofetilide <- build_main("Dofetilide", raw, cqtkit_data_bl_dofetilide)
cqtkit_data_quinidine <- build_main("Quinidine Sulph", raw, cqtkit_data_bl_quinidine)
cqtkit_data_ranolazine <- build_main("Ranolazine", raw, cqtkit_data_bl_ranolazine)
```

The analysis datasets are written with `usethis::use_data()`
