---
title: "Styling plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Styling plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(cqtkit)
```

Every cqtkit plotting function takes a `style` argument. As of 1.2.0 that
argument accepts two things: a `style_spec()` from ggstylekit, or the style
list that `set_style()` builds. The list form is deprecated and will be removed
in 2.0.0.

The two are told apart by class, so an existing call keeps the styling engine
it has always used and produces the figure it has always produced. Nothing
about the list path changed in 1.2.0.

## The spec path

```{r}
eda_scatter_with_regressions(
  cqtkit_data_verapamil,
  deltaQTCF,
  CONC,
  TRTG,
  reference_threshold = c(-10, 10),
  style = style_spec(
    title = "Concentration-QTc Relationship",
    xlabel = "Concentration (ng/mL)",
    ylabel = "deltaQTcF (ms)",
    colors = c("Placebo" = "grey", "Verapamil HCL" = "steelblue")
  )
)
```

cqtkit carries its own defaults, and a spec overrides them field by field.
Anything left `NULL` falls through to the cqtkit default, so a bare
`style_spec()` gives the house figure.

Plots returned from the spec path carry the state ggstylekit needs, so they can
be adjusted afterwards:

```{r}
p <- gof_plots(
  cqtkit_data_verapamil,
  fit,
  deltaQTCF,
  CONC,
  NTLD,
  TRTG,
  style = style_spec()
)

restyle_plot(p, style_spec(legend.position = "bottom"))
reveal(p)
```

Plots returned from the list path do not carry that state, and
`restyle_plot()` and `reveal()` error on them.

## Legends

The legacy vocabulary spells legends as flat arguments. The spec vocabulary
puts them in `legends`, a list of `legend_spec()` entries, one per channel.

```{r}
# deprecated
set_style(legend = "Treatment", color_order = 1, labels = c("Reference -10" = NA))

# current
style_spec(
  legends = list(
    legend_spec(
      channel = "color",
      title = "Treatment",
      order = 1,
      labels = c("Reference -10" = NA)
    )
  )
)
```

A label of `NA` drops that key from the legend, the same idiom in both.

A legend shared across two channels is two `legend_spec()` entries carrying the
same `title`, `labels` and `order`. That is how the unified colour and shape
legend in the `predict_*` plots is built.

`legend_spec(hide = TRUE)` drops a channel's legend outright.

## Argument map

| `set_style()` | `style_spec()` |
| --- | --- |
| `title` | `title` |
| `xlabel` | `xlabel` |
| `ylabel` | `ylabel` |
| `xlims` | `xlims` |
| `ylims` | `ylims` |
| `logx` | `logx` |
| `logy` | `logy` |
| `colors` | `colors` |
| `shapes` | `shapes` |
| `fill_alpha` | `fill_alpha` |
| `legend.position` | `legend.position` |
| `legend.title.position` | `legend.title.position` |
| `legend.title.hjust` | `legend.title.hjust` |
| `legend_nrow` | `legend_nrow` |
| `caption_hjust` | `caption_hjust` |
| `legend` | `legend_spec(channel = "color", title = )` |
| `fill_legend` | `legend_spec(channel = "fill", title = )` |
| `shape_legend` | `legend_spec(channel = "shape", title = )` |
| `labels` | `legend_spec(labels = )` |
| `color_order` | `legend_spec(channel = "color", order = )` |
| `shape_order` | `legend_spec(channel = "shape", order = )` |
| `linetype_order` | `legend_spec(channel = "linetype", order = )` |
| `fill_order` | `legend_spec(channel = "fill", order = )` |

## Do not attach ggstylekit

cqtkit re-exports `style_spec()`, `legend_spec()`, `reveal()` and
`restyle_plot()`, which is the supported way to reach them. Anything else in
ggstylekit should be qualified, as `ggstylekit::series_layer()`.

Do not call `library(ggstylekit)` while both styling APIs exist. Both packages
export `set_style` and `style_plot`, they are unrelated functions, and whichever
package is attached second wins. ggstylekit's `set_style()` takes `(style, ...)`,
so a normal cqtkit call fails with `argument "style" is missing, with no
default`.
