---
title: "Styling plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Styling plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

```{r setup, message = FALSE}
library(cqtkit)
library(ggstylekit)
```

Every cqtkit plotting function returns a `ggplot` object and takes a
`style` argument. Styling is handled by
[ggstylekit](https://github.com/a2-ai/ggstylekit): pass a
`ggstylekit::style_spec()` to control labels, limits, colors, and
legends.

## The `style` argument

`style_spec()` fields cover the common adjustments. Unset fields fall
back to cqtkit's house defaults.

```{r style-basic}
eda_qt_rr_plot(
  cqtkit_data_verapamil, RR, QT, ID, TRTG,
  model_type = "lm",
  style = style_spec(
    xlabel = "RR (ms)",
    ylabel = "QT (ms)",
    colors = c("Placebo" = "grey", "Verapamil HCL" = "steelblue")
  )
)
```

## Legends

Control a legend with `legend_spec()`, naming the aesthetic `channel`
it applies to (`"color"`, `"fill"`, `"shape"`, ...).

```{r style-legend}
eda_qt_rr_plot(
  cqtkit_data_verapamil, RR, QT, ID, TRTG,
  model_type = "lm",
  style = style_spec(
    colors = c("Placebo" = "grey", "Verapamil HCL" = "steelblue"),
    legends = legend_spec(channel = "color", title = "Treatment"),
    legend.position = "bottom"
  )
)
```

## Revealing a covariate

cqtkit plots carry the full dataset, so `ggstylekit::reveal()` can map a
covariate the plot doesn't already show onto a free aesthetic. Name a
channel with `as`, or supply values directly.

```{r reveal}
p <- eda_qt_rr_plot(
  cqtkit_data_verapamil, RR, QT, ID, TRTG,
  model_type = "lm",
  style = style_spec(
    colors = c("Placebo" = "grey", "Verapamil HCL" = "steelblue")
  )
)

reveal(p, SEX, as = "color", on = 'point', colors = c(F = 'pink', M = 'steelblue')) |>
  reveal(SEX, as = 'facet')
```

## Styling model diagnostics

The `style` argument works the same way on goodness-of-fit plots. Fit a
model first, then pass a `style_spec()` to any `gof_*` function.

```{r fit}
fit <- fit_prespecified_model(
  cqtkit_data_verapamil,
  deltaQTCF, ID, CONC, deltaQTCFBL, TRTG, TAFD,
  method = "REML", remove_conc_iiv = TRUE
)
```

`gof_plots()` returns a multi-panel `patchwork`; one `style_spec()`
applies across every panel and the legends are collected.

```{r gof-plots, fig.width = 7, fig.height = 5}
gof_plots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    colors = c("Placebo" = "grey", "Verapamil HCL" = "steelblue"),
    legends = legend_spec(channel = "color", title = "Treatment", order = 1),
    legend.position = "bottom"
  )
) |>
  reveal(SEX, as = "shapes",
         legends = legend_spec(channel = "shape", title = "Sex", order = 2)) |>
  reveal(SEX, as = "fill", on = "bar", fill = c(F = "pink", M = "steelblue"),
         legends = legend_spec(channel = "fill", title = "Sex", order = 3))
```

The same `style_spec()` styles any other `gof_*` function. Concordance
plots compare population and individual predictions:

```{r gof-concordance,  fig.width = 6, fig.height = 7}
gof_concordance_plots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    colors = c("Placebo" = "grey", "Verapamil HCL" = "steelblue"),
    legends = legend_spec(channel = "color", title = "Treatment"),
    legend.position = "bottom",
    facet_nrow = 2
  )
) |>
  reveal(SEX, as = 'facet')
```

Residual boxplots over time map treatment to `fill`, so name that
channel in the legend:

```{r gof-resid-box, fig.width = 7, fig.height = 4}
gof_residuals_time_boxplots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    legends = legend_spec(channel = "fill", title = "Treatment"),
    legend.position = "bottom"
  )
)
```

The VPC plot is a single panel:

```{r gof-vpc}
set.seed(804831)
gof_vpc_plot(cqtkit_data_verapamil, fit, CONC, deltaQTCF, nruns = 10)
```
