---
title: "Styling plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Styling plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
```

```{r setup, message = FALSE}
library(cqtkit)
library(ggstylekit)
```

Every cqtkit plotting function returns a `ggplot` object and takes a
`style` argument handled by
[ggstylekit](https://github.com/a2-ai/ggstylekit). This vignette leans
hard on that integration — it styles and `reveal()`s aggressively so
that each plot exercises a different slice of the styling surface.

## The `style` argument

`style_spec()` fields cover labels, limits, per-series aesthetics, and
the theme all at once.

```{r style-basic}
eda_qt_rr_plot(
  cqtkit_data_verapamil, RR, QT, ID, TRTG,
  model_type = "lm",
  style = style_spec(
    title = "QT vs RR",
    xlabel = "RR interval (ms)",
    ylabel = bquote("QT interval (" * italic(t) * ", ms)"),
    xlims = c(600, 1400),
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    point_alpha = 0.5,
    point_size = 1.5,
    theme = ggplot2::theme_minimal()
  )
) |>
  reveal(SEX, as = c("shapes", "facet"))
```

## Legends

`legend_spec()` controls the title, position, wrapping, and entry
labels of a channel's legend.

```{r style-legend}
eda_qt_rr_plot(
  cqtkit_data_verapamil, RR, QT, ID, TRTG,
  model_type = "lm",
  style = style_spec(
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    legend.position = "bottom",
    legends = legend_spec(
      channel = "color",
      title = "Treatment arm",
      nrow = 1,
      labels = c("Verapamil HCL" = "Verapamil")
    )
  )
)
```

## Revealing a covariate

`reveal()` maps a carried covariate onto a free aesthetic. A
**continuous** covariate becomes a colorbar; a **discrete** one a
keyed legend. Reveals chain.

```{r reveal-continuous}
eda_qt_rr_plot(cqtkit_data_verapamil, RR, QT, ID, model_type = "lm") |>
  reveal(AGE, as = "color")
```

```{r reveal-chained}
eda_qt_rr_plot(
  cqtkit_data_verapamil, RR, QT, ID, TRTG,
  model_type = "lm",
  style = style_spec(colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"))
) |>
  reveal(SEX, as = "shapes", legends = legend_spec(channel = "shape", title = "Sex")) |>
  reveal(RACE, as = "facet")
```

# Exploratory plots

## `eda_qt_rr_plot()`

Continuous size reveal plus a discrete facet, on a log-x axis with
custom breaks.

```{r eda-qt-rr}
eda_qt_rr_plot(
  cqtkit_data_verapamil, RR, QT, ID,
  model_type = "lm",
  style = style_spec(
    logx = TRUE,
    xbreaks = c(700, 900, 1100, 1300),
    point_alpha = 0.4
  )
) |>
  reveal(WGHT, as = "size", legends = legend_spec(title = "Weight (kg)")) |>
  reveal(SEX, as = "facet")
```

## `eda_qtc_comparison_plot()`

A multi-panel patchwork. One `style_spec()` hits every panel, and a
continuous color reveal is grafted across the whole patchwork.

```{r eda-qtc-comparison, fig.width = 8, fig.height = 7}
eda_qtc_comparison_plot(
  cqtkit_data_verapamil, RR, QT, QTCB, QTCF,
  id_col = ID, trt_col = TRTG,
  model_type = "lme", remove_rr_iiv = TRUE,
  style = style_spec(
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    point_alpha = 0.4,
    legend.position = "bottom"
  )
) |>
  reveal(WGHT, as = "color", on = "point",
         legends = legend_spec(title = "Weight (kg)"))
```

## `eda_quantiles_plot()`

Aggregated (not reveal-able), so this leans on style: log axes,
relabeled legend entries, and plotted observations.

```{r eda-quantiles}
eda_quantiles_plot(
  dplyr::filter(cqtkit_data_verapamil, DOSE > 0),
  CONC, QTCF, trt_col = TRTG,
  plot_observations = TRUE,
  style = style_spec(
    logx = TRUE,
    logy = TRUE,
    xlabel = "Concentration (ng/mL, log)",
    ylabel = "QTcF (ms, log)",
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    legends = legend_spec(
      channel = "color", title = "Arm",
      labels = c("Verapamil HCL" = "Verapamil")
    )
  )
)
```

## `eda_scatter_with_regressions()`

Reference lines and regression lines share a `linetype` legend (with
relabeled entries). Color is free, so a discrete covariate is revealed
onto the points, and legend order is set explicitly.

```{r eda-scatter}
eda_scatter_with_regressions(
  cqtkit_data_verapamil, deltaQTCF, CONC, TRTG,
  reference_threshold = c(-10, 10),
  style = style_spec(
    xlabel = "Concentration (ng/mL)",
    ylabel = bquote(Delta ~ "QTcF (ms)"),
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    point_alpha = 0.4,
    legends = list(
      legend_spec(channel = "color", title = "Treatment", order = 1),
      legend_spec(
        channel = "linetype", title = "Regresions", order = 2,
        labels = c("Reference -10" = "+/-10 ms", "Reference 10" = NA)
      )
    )
  )
) |>
  reveal(SEX, as = "shapes", on = "point",
         legends = legend_spec(title = "Sex"))
```

The fit and reference lines are named series (`"LOESS Regression"`,
`"Linear Regression"`, `"Reference <threshold>"`), so the per-series
maps restyle them by name just like data series:

```{r eda-scatter-series}
eda_scatter_with_regressions(
  cqtkit_data_verapamil, deltaQTCF, CONC, TRTG,
  reference_threshold = c(-10, 10),
  style = style_spec(
    colors = c(
      "LOESS Regression" = "firebrick",
      "Linear Regression" = "grey40",
      "Reference 10" = "orange",
      "Reference -10" = "orange"
    ),
    fill = c("LOESS Regression" = "mistyrose"),
    linetypes = c("LOESS Regression" = "solid")
  )
)
```

## `eda_mean_dv_over_time()`

Aggregated with a PK overlay (secondary axis) and error bars. The
reference line merges into the color legend; an entry is relabeled.

```{r eda-mean-dv}
eda_mean_dv_over_time(
  cqtkit_data_verapamil, deltaQTCF, NTLD, DOSEF,
  secondary_data_col = CONC,
  group_col = TRTG, reference_threshold = 10, error_bars = "CI",
  style = style_spec(
    ylabel = bquote("Mean " ~ Delta ~ "QTcF (ms)"),
    xlabel = "Time (h)",
    legend.position = "bottom",
    legend_nrow = 2
  )
)
```

## `eda_hysteresis_loop_plot()`

```{r eda-hysteresis}
p1 <- eda_hysteresis_loop_plot(
  droplevels(dplyr::filter(cqtkit_data_verapamil, DOSE > 0)),
  NTLD, deltaQTCF, CONC, DOSEF,
  style = style_spec(
    ylabel = bquote(Delta ~ "QTcF (ms)"),
    xlabel = "Concentration (ng/mL)",
    theme = ggplot2::theme_minimal(),
    legends = legend_spec(channel = 'color', hide = TRUE)
  )
)

p2 <- eda_hysteresis_loop_plot(
  cqtkit_data_verapamil, 
  NTLD, deltaQTCF, CONC, DOSEF,
  reference_dose = "0 mg",
  style = style_spec(
    ylabel = bquote(Delta ~ Delta ~ "QTcF (ms)"),
    xlabel = "Concentration (ng/mL)",
    theme = ggplot2::theme_minimal(),
    legends = legend_spec(channel = 'color', hide = TRUE)
  )
)

combine_styled_plots(p1, p2, axes = "collect_x")
```

# Goodness-of-fit plots

```{r fit}
fit <- fit_prespecified_model(
  cqtkit_data_verapamil,
  deltaQTCF, ID, CONC, deltaQTCFBL, TRTG, TAFD,
  method = "REML", remove_conc_iiv = TRUE
)
```

## `gof_plots()`

Three legends collected across a 2x2 patchwork: treatment on color, and
SEX revealed onto both point shapes and the histogram fill, with
explicit ordering. The loess smooths here are the same
`"LOESS Regression"` series as in the eda plots, so one
`colors = c("LOESS Regression" = ...)` entry recolors them too.

```{r gof-plots, fig.width = 8, fig.height = 6}
gof_plots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    point_alpha = 0.5,
    legends = legend_spec(channel = "color", title = "Treatment", order = 1),
    legend.position = "right"
  )
) |>
  reveal(SEX, as = "shapes",
         legends = legend_spec(channel = "shape", title = "Sex", order = 2)) |>
  reveal(SEX, as = "fill", on = "bar", fill = c(F = "pink", M = "steelblue"),
         legends = legend_spec(channel = "fill", title = "Sex", order = 3))
```

## `gof_concordance_plots()`

A continuous covariate revealed as a colorbar onto the points, with
facets by SEX laid out in two rows.

```{r gof-concordance, fig.width = 7, fig.height = 7}
gof_concordance_plots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    legend.position = "bottom"
  )
) |>
  reveal(AGE, as = "color", on = "point",
         legends = legend_spec(channel = "color", title = "Age (y)"))
```

## `gof_residuals_plots()`

Discrete covariate revealed onto shapes across the panels.

```{r gof-residuals, fig.width = 8, fig.height = 5}
gof_residuals_plots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    legend.position = "bottom"
  )
) |>
  reveal(SEX, as = "shapes", legends = legend_spec(channel = "shape", title = "Sex"))
```

## `gof_qq_plots()`

```{r gof-qq, fig.width = 7, fig.height = 4}
gof_qq_plots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    colors = c("Placebo" = "grey60", "Verapamil HCL" = "steelblue"),
    legend.position = "bottom"
  )
)
```

## `gof_residuals_time_boxplots()`

Treatment on `fill` with reduced alpha; SEX revealed onto the box
outline color with a thicker line; reference lines keep their own
`linetype` legend.

```{r gof-resid-time-box, fig.width = 8, fig.height = 4.5}
gof_residuals_time_boxplots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    fill = c("Placebo" = "grey70", "Verapamil HCL" = "steelblue"),
    fill_alpha = 0.4,
    legends = legend_spec(channel = "fill", title = "Treatment"),
    legend.position = "right"
  )
) |>
  reveal(SEX, as = "color", on = "box", colors = c(F = "firebrick", M = "navy"),
         box_linewidth = 0.5, legends = legend_spec(channel = "color", title = "Sex"))
```

## `gof_residuals_trt_boxplots()`

```{r gof-resid-trt-box, fig.width = 6, fig.height = 4.5}
gof_residuals_trt_boxplots(
  cqtkit_data_verapamil, fit, deltaQTCF, CONC, NTLD, TRTG,
  style = style_spec(
    fill = c("Placebo" = "grey70", "Verapamil HCL" = "steelblue"),
    fill_alpha = 0.5,
    legends = legend_spec(channel = "fill", title = "Treatment"),
    legend.position = "bottom"
  )
)
```

## `gof_vpc_plot()`

Percentile lines own the color channel, so SEX is revealed onto shapes.

```{r gof-vpc}
set.seed(804831)
gof_vpc_plot(cqtkit_data_verapamil, fit, CONC, deltaQTCF, nruns = 10) |>
  reveal(SEX, as = "shapes", on = "point",
         legends = legend_spec(channel = "shape", title = "Sex"))
```

# Exposure-response predictions

## `predict_with_observations_plot()`

The observations carry the full dataset, so a covariate can be revealed
onto them. This holds for both the single-treatment (ΔQTcF) form and the
contrast (ΔΔQTcF) form built with `control_predictors`.

Single treatment:

```{r predict-observations}
predict_with_observations_plot(
  cqtkit_data_verapamil, fit, CONC, deltaQTCF,
  treatment_predictors = list(
    CONC = 0, deltaQTCFBL = 0, TRTG = "Verapamil HCL", TAFD = "0.5 HR"
  ),
  reference_threshold = 10,
  style = style_spec(ylabel = bquote(Delta ~ "QTcF (ms)"))
) |>
  reveal(SEX, as = "shapes", on = "point",
         legends = legend_spec(channel = "shape", title = "Sex"))
```

Placebo-corrected contrast (ΔΔQTcF), revealing the same covariate on the
contrast observations:

```{r predict-observations-contrast}
predict_with_observations_plot(
  cqtkit_data_verapamil, fit, CONC, deltaQTCF,
  id_col = ID, ntime_col = NTLD, trt_col = TRTG,
  treatment_predictors = list(
    CONC = 0, deltaQTCFBL = 0, TRTG = "Verapamil HCL", TAFD = "2 HR"
  ),
  control_predictors = list(
    CONC = 0, deltaQTCFBL = 0, TRTG = "Placebo", TAFD = "2 HR"
  ),
  reference_threshold = c(-10, 10),
  style = style_spec(ylabel = bquote(Delta ~ Delta ~ "QTcF (ms)"))
) |>
  reveal(SEX, as = "shapes", on = "point",
         legends = legend_spec(channel = "shape", title = "Sex"))
```

## `predict_with_quantiles_plot()`

```{r predict-quantiles}
predict_with_quantiles_plot(
  cqtkit_data_verapamil, fit, CONC, deltaQTCF,
  treatment_predictors = list(
    CONC = 0, deltaQTCFBL = 0, TRTG = "Verapamil HCL", TAFD = "0.5 HR"
  ),
  reference_threshold = 10, error_bars = "CI",
  style = style_spec(
    ylabel = bquote(Delta ~ "QTcF (ms)"),
    xlabel = "Concentration (ng/mL)"
  )
)
```

## `predict_with_exposure_plot()`

`cmaxes` marks exposure levels via the geometric-mean Cmax.

```{r predict-exposure}
pk_df <- compute_pk_parameters(
  dplyr::filter(cqtkit_data_verapamil, DOSE != 0), ID, DOSEF, CONC, NTLD
)

predict_with_exposure_plot(
  cqtkit_data_verapamil, fit, CONC,
  treatment_predictors = list(
    CONC = 0, deltaQTCFBL = 0, TRTG = "Verapamil HCL", TAFD = "0.5 HR"
  ),
  cmaxes = pk_df[[1, "Cmax_gm"]],
  reference_threshold = 10,
  style = style_spec(ylabel = bquote(Delta ~ "QTcF (ms)"))
)
```
